---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# hrd

<!-- badges: start -->
<!-- badges: end -->

`hrd` is a collection of functions to help implement, develop and fit Hüsler-Reiss distribution based models.

## Installation

You can install the development version of hrd like so:

``` r
install.packages('devtools')
devtools::install_github("dairer/hrd")
```

```{r results='hide'}
library(hrd)
library(ggplot2) 
library(tidyverse)
```

## Example
### Sampling bivariate Hüsler-Reiss copula 
```{r example1}
# sample data from a bivariate Hüsler-Reiss copula
data = rhr(n=1000, lambda = 1.5)

# visualise samples
data %>%
  as.data.frame() %>%
  ggplot()+
  geom_point(aes(u,v), shape = 4)+
  theme_minimal(12)
```

### Esimating the dependence parameter of a Hüsler-Reiss copula
```{r example2}
fit_hrc(u = data[,1], v = data[,2], initial_est = 1)
```
### Esimating generalised pareto marings and dependence parameter of a Hüsler-Reiss copula jointly
```{r example3}
# transform margins to be GP distributed
x = qgp(data[,1], scale = 2, shape = -0.1)
y = qgp(data[,2], scale = 1.3, shape = -0.1)

# visualise the multivariate distribution
data.frame(x,y) %>%
  ggplot()+
  geom_point(aes(x,y), shape = 4)+
  theme_minimal()

model_fit = fit_hrc_gp(x,y,initial_est = c(1,0,1,0,1))

# visualise estimates
data.frame(actual = c(2, -0.1, 1.3, -0.1, 1.5),
       estimates = model_fit$estimate,
       lower = model_fit$ci[,1],
       upper = model_fit$ci[,2],
       params = c('σ1', 'ξ1', 'σ2', 'ξ2', 'λ')) %>%
  ggplot()+
  geom_point(aes(params, estimates))+
  geom_point(aes(params, actual),position = position_nudge(x = -0.075), shape = 4, col = 'red')+
  geom_segment(aes(x = params, xend = params, y = lower, yend = upper))+
  labs(x = "Parameters",
       y = "Parameter estimates")+
  theme_minimal(12)
```

In each case the red 'x' is the true value. The black point along with the segment are the point estimates and 95% confidence intervals.
λ is the dependence parameter of the Hüsler-Reiss copula.
σ1 and ξ1 are the scale and shape parameters of GPD margin 1. Equivalent for margin 2.



### We can compare estimates from margin-dependence indpendent analysis (in magenta) to joint modelling (in black).
```{r example4}
dependence_mod = fit_hrc(data[,1], data[,2], initial_est = 1)

marg1_mod = fit_gpd(x, initial_est = c(1,0))
marg2_mod = fit_gpd(y, initial_est = c(1,0))

data.frame(actual = c(2, -0.1, 1.3, -0.1, 1.5),
       estimates_joint = model_fit$estimate,
       lower_joint = model_fit$ci[,1],
       upper_joint = model_fit$ci[,2],
       estimates_indep = c(marg1_mod$estimate, marg2_mod$estimate, dependence_mod$estimate),
       lower_indep = c(marg1_mod$ci[,1],marg2_mod$ci[,1],dependence_mod$ci[1]),
       upper_indep = c(marg1_mod$ci[,2],marg2_mod$ci[,2],dependence_mod$ci[2]),
       params = c('σ1', 'ξ1', 'σ2', 'ξ2', 'λ')) %>%
  ggplot()+
  geom_point(aes(params, actual),position = position_nudge(x = -0.075), shape = 4, col = 'red')+
  geom_point(aes(params, estimates_joint))+
  geom_segment(aes(x = params, xend = params, y = lower_joint, yend = upper_joint))+
  geom_point(aes(params, estimates_indep),position = position_nudge(x = 0.075), col = "magenta")+
  geom_segment(aes(x = params, xend = params, y = lower_indep, yend = upper_indep),position = position_nudge(x = 0.075), col = "magenta")+
  labs(x = "Parameters",
       y = "Parameter estimates")+
  theme_minimal(12)
```


### We can fit a bi-variate copula using a bayesian framework.
```{r example5}
data = rhr(n=1000, lambda = 1.5)
my_bayesian_fit = fit_hrc_bay(data[,1], 
                              data[,2],
                              chains = 2, 
                              prior_mean = 1, # mean of normal prior
                              prior_sd = 2,
                              cores = 2, # !! Make sure you have enough cores !!
                              iter = 1000,
                              thin = 2)


data.frame(lambda_post = my_bayesian_fit$post,
           lambda = my_bayesian_fit$estimate,
           actual = 1.5) %>%
  ggplot()+
  geom_density(aes(lambda_post), alpha = 0.5, col = 'black', fill = 'lightblue')+
  geom_vline(aes(xintercept = actual), col = 'black', size = 1.5)+
  theme_minimal()+
  labs(x = "λ",
       y = "Density")
```

The function `fit_hrc_bay` retuns the fitted model so you can do all the normal stan things, e.g. investigate trace plots:

```{r ex5}
rstan::traceplot(my_bayesian_fit$fitted_model)
```
