---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# hrd

<!-- badges: start -->
<!-- badges: end -->

`hrd` is a collection of functions to help implement and develop and fit Hüsler-Reiss models.

## Installation

You can install the development version of hrd like so:

``` r
install.packages('devtools')
devtools::install_github("dairer/hrd")
```

## Example

### Sampling bivariate Hüsler-Reiss copula 
```{r example1}
library(hrd)
library(ggplot2)

# sample data from a bivariate Hüsler-Reiss copula
data = rhr(n=500, lambda = 1.5)

# visualise samples
data %>%
  as.data.frame() %>%
  ggplot()+
  geom_point(aes(u,v), shape = 4)+
  theme_minimal(12)
```

### Esimating the dependence parameter of a Hüsler-Reiss copula
```{r example2}
fit_hrc(u = data[,1], v = data[,2], initial_est = 1)
```
### Esimating generalised pareto marings and dependence parameter of a Hüsler-Reiss copula jointly
```{r example3}
# transform margins to be GP distributed
x = qgp(data[,1], scale = 2, shape = -0.1)
y = qgp(data[,2], scale = 1.3, shape = -0.1)

# visualise the multivariate distribution
data.frame(x,y) %>%
  ggplot()+
  geom_point(aes(x,y), shape = 4)+
  theme_minimal()

fit_hrc_gp(x,y,initial_est = c(1,0,1,0,1))
```

